---
title: "Exercises 1"
output: github_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Visualization: Flights at ABIA

```{r abia_background, include=FALSE}

library(Hmisc)
library(tidyverse)
library(FNN)
library(mosaic)
library(maps)
library(mapdata)

#read data
abia = read.csv('ABIA.csv')

#clean & subset
abia_cr<- abia
abia_cr$DepDelay[abia_cr$Cancelled == 1] <- 10000

abia_crd <- subset(abia_cr, Dest != 'AUS')
abia_crd$LongWait <- (abia_crd$DepDelay >= 20)

abia_sumbydest = abia_crd  %>%
  group_by(Dest) %>%
  summarize(median.DepDelay = median(DepDelay), FlightCount = length(Dest), Prob_LW = mean(LongWait))

#Filter groups with not enough data
abia_sumbydest <- subset(abia_sumbydest, FlightCount >= 50)
abia_sumbydest$Prob_LW = abia_sumbydest$Prob_LW * 100

#Import Airport codes
codes = read.csv('airport-codes.csv')
codes = separate(codes,'coordinates', c('Lat','Long') , sep=',' , remove=TRUE, convert=TRUE)

#filter codes & clean
codes_f = codes[codes$iata_code %in% abia_sumbydest$Dest,]
codes_f = separate(codes_f,'iso_region', c('CountryCode','State') , sep='-' , remove=TRUE, convert=TRUE)
codes_f$Dest = codes_f$iata_code

#merge to grouped data
abia_merged = merge(abia_sumbydest, codes_f, by='Dest')

##map data
usa <- map_data('usa')
states <- map_data('state')

```

```{r abia_plots, echo = FALSE}

ggUSA <- 
  ggplot() + 
  geom_polygon(data = states, aes(x=long, y = lat, group = group), color = 'white', fill = 'light blue') + 
  coord_fixed(1.3) 

ggUSA + 
  geom_point(data = abia_merged, aes(x = Long, y = Lat, fill = Prob_LW), shape = 23, size = 3, stroke = 1.75, alpha = .75) +
  geom_text(data = abia_merged, aes(x = Long, y = Lat, label = abia_merged$Dest ), check_overlap = TRUE, nudge_x = -2, size = 4) + 
  scale_fill_gradient(low = 'green', high = 'red',guide = 'colourbar') +
  guides(fill = guide_colorbar(title = '')) +
  ggtitle('Departures - Percent Chance of a Delay Greater Than 20 Minutes by Destination') + 
  theme(
    axis.text = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    panel.border = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_blank(),
    axis.title = element_blank(),
    plot.title = element_text(hjust = 0)
  ) 

```


## K-Nearest Neighbors Model: Car Price


## Below includes basic template instructions 

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
